### 2022-05-23 assets/tag_distribution.tex
:paintbrush: Illustrate probability distribution

\begin{tikzpicture}
    \begin{groupplot}
        [group style = {%
        rows = 3,
        group name = plots,
%         xlabels at = edge bottom,
        x descriptions at = edge bottom,
        y descriptions at = edge left,
        vertical sep = 15
        },
        ybar,
        xmin = 0,
        xmax = 4,
        xtick = data,
        ymin = 0,
        ymax = 1,
        xlabel={Reflection State},
        ylabel={Probability},
        enlarge x limits = {value = 0.25, upper},
        enlarge y limits = {value = 0.1, upper},
        width = 10cm,
        height = 4cm
        ]
        \nextgroupplot
        \addplot[blue,fill=blue!30!white] coordinates {(1,0.25) (2,0.25) (3,0.25) (4,0.25)};
        \legend{SR}
        \nextgroupplot
        \addplot[red,fill=red!30!white] coordinates {(1,0) (2,1) (3,0) (4,0)};
        \legend{IRS}
        \nextgroupplot
        \addplot[brown!60!black!,fill=brown!30!white] coordinates {(1,0.2) (2,0.4) (3,0.3) (4,0.1)};
        \legend{Metascatter}
    \end{groupplot}
\end{tikzpicture}