### 2022-05-26 assets/scatter_model.tex

\begin{circuitikz}[transform shape]
	\tikzstyle{every node}=[font=\Large]
	\draw (0,0) node[bareantenna](bareantenna){};
	\draw (bareantenna.west) ++(-1.5,0) node[waves](WI){};
	\draw (WI.north east) ++(0.25,0) node[font=\Large]{$\vec{E}_{\mathrm{I}}$};
	\draw (WI.south east) ++(0.25,0) node[font=\Large]{$\vec{H}_{\mathrm{I}}$};
	\draw (bareantenna.east) ++(0.875,0) node[waves](WR){};
	\draw (WR.north east) ++(0.35,0) node[font=\Large]{$\vec{E}_m$};
	\draw (WR.south east) ++(0.35,0) node[font=\Large]{$\vec{H}_m$};
	\draw (bareantenna)
	to [R=$Z_{\mathrm{A}}$] ++(0,-1.5) node[rotary switch=4 in 90 wiper 30,anchor=ext center,rotate=270](SW){};
	\draw (SW.cout 2) node[below=0.1cm,font=\Large]{$m$}
	to [R=$Z_m$] ++(2,0) node[ground]{};
\end{circuitikz}