### 2022-06-06 assets/scatter_model.tex
:paintbrush: Update scatter model plot

% \begin{circuitikz}[transform shape]
% 	\tikzstyle{every node}=[font=\Large]
% 	\draw (0,0) node[bareantenna](bareantenna){};
% 	\draw (bareantenna.west) ++(-1.5,0) node[waves](WI){};
% 	\draw (WI.north east) ++(0.25,0) node[font=\Large]{$\vec{E}_{\mathrm{I}}$};
% 	\draw (WI.south east) ++(0.25,0) node[font=\Large]{$\vec{H}_{\mathrm{I}}$};
% 	\draw (bareantenna.east) ++(0.875,0) node[waves](WR){};
% 	\draw (WR.north east) ++(0.35,0) node[font=\Large]{$\vec{E}_m$};
% 	\draw (WR.south east) ++(0.35,0) node[font=\Large]{$\vec{H}_m$};
% 	\draw (bareantenna)
% 	to [R=$Z_{\mathrm{A}}$] ++(0,-1.5) node[rotary switch=4 in 90 wiper 30,anchor=ext center,rotate=270](SW){};
% 	\draw (SW.cout 2) node[below=0.1cm,font=\Large]{$m$}
% 	to [R=$Z_m$] ++(2,0) node[ground]{};
% \end{circuitikz}

\begin{circuitikz}[transform shape]
	\tikzstyle{every node}=[font=\large]
	\draw (0,0) node[bareantenna](bareantenna){};
	\draw (bareantenna.west) ++(-1.5,0) node[waves](WI){};
	\draw (WI.north east) ++(0.25,0) node{$\vec{E}_{\mathrm{I}}$};
	\draw (WI.south east) ++(0.25,0) node{$\vec{H}_{\mathrm{I}}$};
	\draw (bareantenna.east) ++(0.875,0) node[waves](WR){};
	\draw (WR.north east) ++(0.35,0) node{$\vec{E}_m$};
	\draw (WR.south east) ++(0.35,0) node{$\vec{H}_m$};
	\draw (bareantenna) to ++(0,-1.5) to [generic,l=$Z_{\mathrm{A}}$,/tikz/circuitikz/bipoles/length=0.75cm] ++(3.5,0) node[rotary switch=4 in 90 wiper 30,anchor=ext center,rotate=0](SW){};
	\draw (SW.cout 1) to ++(0.625,+0.5) to [generic,/tikz/circuitikz/bipoles/length=0.75cm] ++(3,0) coordinate(E1);
	\draw (SW.cout 2) node[above right]{$m$} to [generic,l=$Z_m$,/tikz/circuitikz/bipoles/length=0.75cm] ++(3,0) coordinate(E2);
	\draw (SW.cout 3) to [generic,/tikz/circuitikz/bipoles/length=0.75cm] ++(3,0) coordinate(E3);
	\draw (SW.cout 4) to ++(0.625,-0.5) to [generic,/tikz/circuitikz/bipoles/length=0.75cm] ++(3,0) coordinate(E4);
	% \coordinate (E) at ($(E2)! 0.5!(E3)$);
	\draw (E1) to[short,-*] (E2);
	\draw (E2) to[short,-*] (E3);
	% \draw (E2) to[short,-*] (E);
	% \draw (E) to[short,-*] (E3);
	\draw (E3) to[short] (E4);
	\draw (E2) to[short] ++(1,0) node[ground]{};
\end{circuitikz}