### 2022-06-08 assets/trend_distribution.tex
:bento: Update assets

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor3}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor4}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=4.036in,
height=3.396in,
at={(0.677in,0.458in)},
scale only axis,
xmin=1,
xmax=4,
xtick={1, 2, 3, 4},
xlabel style={font=\color{white!15!black}},
xlabel={Reflection State},
ymin=0,
ymax=1,
ylabel style={font=\color{white!15!black}},
ylabel={Probability Distribution},
axis background/.style={fill=white},
xminorgrids,
yminorgrids,
legend style={legend cell align=left, align=left, draw=white!15!black},
title style={font=\huge}, label style={font=\huge}, ticklabel style={font=\LARGE}, legend style={font=\LARGE}
]
\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
1	0.249999735242584\\
2	0.250000264588401\\
3	0.250000265053761\\
4	0.249999735115254\\
};
\addlegendentry{$\rho =0$}

\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
1	0.295743068427155\\
2	0.375240201334651\\
3	0.156312315031268\\
4	0.172704415206926\\
};
\addlegendentry{$\rho =0.002$}

\addplot [color=mycolor3, dotted, line width=2.0pt, mark=square, mark options={solid, mycolor3}]
  table[row sep=crcr]{%
1	0.261901350518658\\
2	0.712510408293707\\
3	0.00842749500731859\\
4	0.0171607461803161\\
};
\addlegendentry{$\rho =0.005$}

\addplot [color=mycolor4, dashdotted, line width=2.0pt, mark=x, mark options={solid, mycolor4}]
  table[row sep=crcr]{%
1	0.139913707035647\\
2	0.859278410930712\\
3	0.000142053828609263\\
4	0.000665828205031897\\
};
\addlegendentry{$\rho =0.01$}

\addplot [color=mycolor5, line width=2.0pt, mark=triangle, mark options={solid, mycolor5}]
  table[row sep=crcr]{%
1	1.29237706072384e-05\\
2	0.999987076229393\\
3	4.46101240540389e-26\\
4	3.49723978579285e-21\\
};
\addlegendentry{$\rho =1$}

\end{axis}
\end{tikzpicture}%