### 2022-06-21 assets/distribution_weights.tex
:bento: Update assets

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor3}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor4}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=4.036in,
height=3.396in,
at={(0.677in,0.458in)},
scale only axis,
xmin=1,
xmax=4,
xtick={1, 2, 3, 4},
xlabel style={font=\color{white!15!black}},
xlabel={Reflection State},
ymin=0,
ymax=1,
ylabel style={font=\color{white!15!black}},
ylabel={Probability Distribution},
axis background/.style={fill=white},
xminorgrids,
yminorgrids,
legend style={legend cell align=left, align=left, draw=white!15!black},
title style={font=\huge}, label style={font=\huge}, ticklabel style={font=\LARGE}, legend style={font=\LARGE}
]
\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
1	2.62331519807269e-06\\
2	0.999997376684802\\
3	4.89949641833056e-30\\
4	2.6174774451771e-24\\
};
\addlegendentry{$\rho =1$}

\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
1	9.21083675347438e-06\\
2	0.812542727242104\\
3	0.187445044005617\\
4	3.01791552545155e-06\\
};
\addlegendentry{$\rho =0.15$}

\addplot [color=mycolor3, dotted, line width=2.0pt, mark=square, mark options={solid, mycolor3}]
  table[row sep=crcr]{%
1	1.56711271328569e-05\\
2	0.695664383940126\\
3	0.304312342609931\\
4	7.60232280996882e-06\\
};
\addlegendentry{$\rho =0.1$}

\addplot [color=mycolor4, dashdotted, line width=2.0pt, mark=x, mark options={solid, mycolor4}]
  table[row sep=crcr]{%
1	1.79207354228568e-05\\
2	0.591606048367841\\
3	0.408364572353122\\
4	1.14585436146039e-05\\
};
\addlegendentry{$\rho =0.05$}

\addplot [color=mycolor5, line width=2.0pt, mark=triangle, mark options={solid, mycolor5}]
  table[row sep=crcr]{%
1	2.61368383798461e-05\\
2	0.497964372118738\\
3	0.501988384568931\\
4	2.11064739508265e-05\\
};
\addlegendentry{$\rho =0$}

\end{axis}
\end{tikzpicture}%