### 2022-07-16 assets/energy_distribution.tex
:paintbrush: Update illustrations

\begin{tikzpicture}
	\begin{axis}[%
			height=5cm,
			width=9cm,
			xlabel = $z$,
			ylabel = {Probability Density},
			font=\scriptsize,
			no markers,
			xmin=0,
			ymin=0,
			xmax=70,
			%     ticks=none,
			xtick={0,12.78,20.28,29.62,70},
			xticklabels={$t_0$,$t_1$,$t_2$,$t_3$,$t_4$},
			yticklabels={},
			samples = 200]

		\addplot+[thick] gnuplot[raw gnuplot] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,1)};
		\addlegendentryexpanded{$f(z \mid \mathcal{H}_1)$}

		\addplot+[thick] gnuplot[raw gnuplot] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.6)};
		\addlegendentryexpanded{$f(z \mid \mathcal{H}_2)$}

		\addplot+[thick,color=brown] gnuplot[raw gnuplot] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.4)};
		\addlegendentryexpanded{$f(z \mid \mathcal{H}_3)$}

		\addplot+[thick,color=green!40!gray] gnuplot[raw gnuplot] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.28)};
		\addlegendentryexpanded{$f(z \mid \mathcal{H}_4)$}
		\draw[color=gray,thick,dashed] (12.78,0) -- (12.78,0.1);
		\draw[color=gray,thick,dashed] (20.28,0) -- (20.28,0.1);
		\draw[color=gray,thick,dashed] (29.62,0) -- (29.62,0.1);
		\draw[latex-latex] (0,0.05) -- (12.78,0.05) node[below,midway]{$\mathcal{R}(\mathcal{H}_1)$};
		\draw[latex-latex] (12.78,0.05) -- (20.28,0.05) node[below,midway]{$\mathcal{R}(\mathcal{H}_2)$};
		\draw[latex-latex] (20.28,0.05) -- (29.62,0.05) node[below,midway]{$\mathcal{R}(\mathcal{H}_3)$};
		\draw[latex-latex] (29.62,0.05) -- (70,0.05) node[below,midway]{$\mathcal{R}(\mathcal{H}_4)$};
	\end{axis}
\end{tikzpicture}