### 2022-07-18 assets/time_block_structure.tex
:bento: Update assets

\begin{tikzpicture}

	% Primary block
	\foreach \x in {0,...,31}{
			\draw (0.25*\x,3) rectangle ++(0.25,0.25);
		}
	\node[font=\LARGE] at (-0.75,3.125) {$\cdots$};
	\node[font=\LARGE] at (8.75,3.125) {$\cdots$};
	\node[font=\LARGE] at (-3.75,3.125) {Primary Block};

	\draw (10.125,3) rectangle ++(0.25,0.25);
	\node[font=\LARGE,right] at (11,3.125){$T_s$};

	% Backscatter block
	\foreach \x in {0,...,7}{
			\draw (\x,1.75) rectangle ++(1,0.5);
		}
	\node[font=\LARGE] at (-0.75,2) {$\cdots$};
	\node[font=\LARGE] at (8.75,2) {$\cdots$};
	\node[font=\LARGE] at (-3.75,2) {Backscatter Block};

	\draw (10,1.875) rectangle ++(0.5,0.25);
	\node[font=\LARGE,right] at (11,2){$T_c$};

	% SR
	\draw [blue,fill=blue!30!white,postaction={pattern=north west lines},pattern color=.](0,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=vertical lines},pattern color=.](1,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=horizontal lines},pattern color=.](2,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=north east lines},pattern color=.](3,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=north east lines},pattern color=.](4,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=vertical lines},pattern color=.](5,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=horizontal lines},pattern color=.](6,0) rectangle ++(1,1);
	\draw [blue,fill=blue!30!white,postaction={pattern=north west lines},pattern color=.](7,0) rectangle ++(1,1);
	\node[font=\LARGE] at (-0.75,0.5) {$\cdots$};
	\node[font=\LARGE] at (8.75,0.5) {$\cdots$};
	\node[font=\LARGE] at (-3.75,0.5) {SR};

	% IRS
	\foreach \x in {0,...,7}{
			\draw [red,fill=red!30!white,postaction={pattern=north west lines},pattern color=.](\x,-1.5) rectangle ++(1,1);
		}
	\node[font=\LARGE] at (-0.75,-1) {$\cdots$};
	\node[font=\LARGE] at (8.75,-1) {$\cdots$};
	\node[font=\LARGE] at (-3.75,-1) {IRS};

	% SR
	\draw [brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.](0,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=vertical lines},pattern color=.](1,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.](2,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.](3,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.](4,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=vertical lines},pattern color=.](5,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=horizontal lines},pattern color=.](6,-3) rectangle ++(1,1);
	\draw [brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.](7,-3) rectangle ++(1,1);
	\node[font=\LARGE] at (-0.75,-2.5) {$\cdots$};
	\node[font=\LARGE] at (8.75,-2.5) {$\cdots$};
	\node[font=\LARGE] at (-3.75,-2.5) {Metascatter};

	\draw [pattern=horizontal lines](10,0.5) rectangle ++(0.5,0.5);
	\draw [pattern=north west lines](10,-2/3) rectangle ++(0.5,0.5);
	\draw [pattern=vertical lines](10,-11/6) rectangle ++(0.5,0.5);
	\draw [pattern=north east lines](10,-3) rectangle ++(0.5,0.5);
	\node[font=\LARGE,right] at (11,0.75){State 1};
	\node[font=\LARGE,right] at (11,-5/12){State 2};
	\node[font=\LARGE,right] at (11,-19/12){State 3};
	\node[font=\LARGE,right] at (11,-2.75){State 4};

	\draw (-6.375,-3.5) rectangle ++(20, 7.375);

\end{tikzpicture}