### 2022-07-22 assets/simulation/distribution_weights_1.tex
:bento: Update input distribution vs weights

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor3}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor4}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=3.907in,
height=3.328in,
at={(0.655in,0.449in)},
scale only axis,
xmin=1,
xmax=4,
xtick={1, 2, 3, 4},
xlabel style={font=\color{white!15!black}},
xlabel={Reflection State},
ymin=0,
ymax=1,
ylabel style={font=\color{white!15!black}},
ylabel={Probability Distribution},
axis background/.style={fill=white},
xmajorgrids,
xminorgrids,
ymajorgrids,
yminorgrids,
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black},
title style={font=\huge}, label style={font=\huge}, ticklabel style={font=\LARGE}, legend style={font=\LARGE}
]
\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
1	1.34887442178325e-06\\
2	0.401147907621958\\
3	0.346863819336987\\
4	0.251986924166634\\
};
\addlegendentry{$\rho =0$}

\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
1	2.43160039284684e-06\\
2	0.373071655693474\\
3	0.387092084568394\\
4	0.239833828137739\\
};
\addlegendentry{$\rho =0.1$}

\addplot [color=mycolor3, dotted, line width=2.0pt, mark=square, mark options={solid, mycolor3}]
  table[row sep=crcr]{%
1	1.27356787994262e-05\\
2	0.337563209241624\\
3	0.428718130958926\\
4	0.23370592412065\\
};
\addlegendentry{$\rho =0.2$}

\addplot [color=mycolor4, dashdotted, line width=2.0pt, mark=x, mark options={solid, mycolor4}]
  table[row sep=crcr]{%
1	2.82495749022891e-05\\
2	0.293114805811411\\
3	0.479872318504251\\
4	0.226984626109436\\
};
\addlegendentry{$\rho =0.3$}

\addplot [color=mycolor5, line width=2.0pt, mark=triangle, mark options={solid, mycolor5}]
  table[row sep=crcr]{%
1	6.85130791403468e-06\\
2	8.96232761727522e-19\\
3	0.999993126219272\\
4	2.24728139461691e-08\\
};
\addlegendentry{$\rho =1$}

\end{axis}
\end{tikzpicture}%