### 2022-08-30 assets/illustration/ambient_backscatter_communication.tex
:bento: Update illustration assets

\begin{tikzpicture}[every node/.style={draw}]
	\node at (0,0) [trapezium,minimum height=0.75cm,anchor=center] {};
	\node at (3,4.5) [isosceles triangle,isosceles triangle stretches,shape border rotate=+270,minimum size=1cm,minimum height=1cm,anchor=north] {};
	\node at (7,4) [rectangle,minimum size=1cm] {};
	\node at (10,0) [trapezium,rotate=180,minimum height=0.75cm,anchor=center] {};

	\node[draw=none] at (0,-0.75) {Legacy Transmitter};
	\node[draw=none] at (3,5) {Backscatter Reader};
	\node[draw=none] at (7,5) {Scattering Node};
	\node[draw=none] at (10,-0.75) {Legacy Receiver};

	\draw[loosely dashed,-latex] (1,0) -- (9,0);
	\draw[loosely dashed,-latex] (0,0.75) -- (2.5,4);
	\draw[loosely dashed,-latex] (0.75,0.5) -- (6.25,3.5);
	\draw[dashdotted,-latex] (6.25,4.5) -- (3.75,4.5);
	\draw[dashdotted,-latex] (7.75,4) -- (10,0.75);
\end{tikzpicture}