### 2022-08-30 assets/illustration/bistatic_backscatter_communication.tex

\begin{tikzpicture}[every node/.style={draw}]
	\node at (0,0) [isosceles triangle,isosceles triangle stretches,shape border rotate=+90,minimum size=1cm,minimum height=1cm,anchor=north] {};
	\node at (7,4) [rectangle,minimum size=1cm] {};
	\node at (10,0) [isosceles triangle,isosceles triangle stretches,shape border rotate=+270,minimum size=1cm,minimum height=1cm,anchor=north] {};

	\node[draw=none] at (0,-1.5) {Carrier Emitter};
	\node[draw=none] at (7,5) {Scattering Node};
	\node[draw=none] at (10,-1.5) {Backscatter Reader};

	\draw[-latex] (1,-1) -- (9.5,-1);
	\draw[-latex] (0.5,0) -- (6,4);
	\draw[loosely dotted,-latex] (8,4) -- (10,0.5);
\end{tikzpicture}