### 2022-08-30 assets/illustration/reconfigurable_intelligent_surface.tex

\begin{tikzpicture}[every node/.style={draw}]
	\node at (0,0) [trapezium,minimum height=0.75cm,anchor=center] {};
	\node at (7,4) [rectangle,minimum size=1cm] {};
	\node at (10,0) [trapezium,rotate=180,minimum height=0.75cm,anchor=center] {};

	\node[draw=none] at (0,-0.75) {Transmitter};
	\node[draw=none] at (7,5) {Reflecting Element};
	\node[draw=none] at (10,-0.75) {Receiver};

	\draw[loosely dashed,-latex] (1,0) -- (9,0);
	\draw[loosely dashed,-latex] (0.75,0.5) -- (6.25,4);
	\draw[loosely dashed,-latex] (7.75,4) -- (10,0.75);
\end{tikzpicture}