### 2022-08-31 assets/illustration/ambient_backscatter_communication.tex
:bento: Update scattering paradigms

\begin{tikzpicture}[font=\LARGE,every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [dart,shape border rotate=90,minimum width=1cm] {};
	\node at (1.5,3.5) [isosceles triangle,isosceles triangle stretches,shape border rotate=+270,minimum size=1cm,minimum height=1cm,anchor=north] {};
	\node at (6.5,3) [rectangle,minimum size=1cm] {};
	\node at (8,0.35) [dart,shape border rotate=270,minimum width=1cm] {};

	\node[draw=none] at (0,-1.375) {Legacy\\Transmitter};
	\node[draw=none] at (1,4.375) {Backscatter\\Reader};
	\node[draw=none] at (7,4.375) {Scattering\\Node};
	\node[draw=none] at (8,-1.375) {Legacy\\Receiver};

	\draw[decorate,decoration={waves,segment length=5mm}] (1,-0.5) -- (7.5,-0.5);
	\draw[decorate,decoration={waves,segment length=5mm}] (0.25,1) -- (0.75,3);
	\draw[decorate,decoration={waves,segment length=5mm}] (0.75,0.25) -- (5.5,2.5);
	\draw[decorate,decoration={triangles,segment length=2.5mm}] (5.5,3.5) -- (2.5,3.5);
	\draw[decorate,decoration={triangles,segment length=2.5mm}] (7.5,3) -- (8,1);

	\node[draw=none] at (4,0.35) {Modulated\\Signal};
	\node[draw=none] at (4,3.5) {Superimposed\\Signal};
\end{tikzpicture}