### 2022-08-31 assets/illustration/riscatter.tex

\begin{tikzpicture}[font=\LARGE,every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [circular sector,shape border rotate=90,minimum width=2cm] {};
	\node at (5,3) [circle,minimum size=1.25cm] {};
	\node at (8,0.55) [circular sector,shape border rotate=270,minimum width=2cm] {};

	\node[draw=none] at (0,-1.25) {Primary\\Transmitter};
	\node[draw=none] at (5,4.5) {RIScatter\\Node};
	\node[draw=none] at (8,-1.25) {Cooperative\\Receiver};

	\draw[decorate,decoration={waves,segment length=5mm}] (1,-0.25) -- (7.375,-0.25);
	\draw[decorate,decoration={waves,segment length=5mm}] (0.5,0.75) -- (4,3);
	\draw[decorate,decoration={triangles,segment length=2.5mm}] (6,3) -- (8,1.25);

	\node[draw=none] at (4,0.875) {Modulated\\Signal};
	\node[draw=none] at (7.8,2.9) {Superimposed\\Signal};
\end{tikzpicture}