### 2022-10-13 assets/illustration/bbc.tex

\begin{tikzpicture}[font=\LARGE,every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [isosceles triangle,isosceles triangle stretches,shape border rotate=+90,minimum size=1cm,minimum height=1cm,anchor=north] {};
	\node at (5,2) [rectangle,minimum size=1cm] {};
	\node at (8,0) [isosceles triangle,isosceles triangle stretches,shape border rotate=+270,minimum size=1cm,minimum height=1cm,anchor=north] {};

	\node[draw=none] at (0,-1.95) {Carrier\\Emitter};
	\node[draw=none] at (5,3.35) {Scattering\\Node};
	\node[draw=none] at (8,-1.95) {Backscatter\\Reader};

	\draw[decorate,decoration={snake,post length=2.5mm},-latex] (1,-1) -- (7.5,-1);
	\draw[decorate,decoration={snake,post length=2.5mm},-latex] (0.5,0) -- (4,2);
	\draw[decorate,decoration={snake,pre length=7.5mm,post length=7.5mm},-latex] (6,2) -- (8,0.35);

	\node[draw=none] at (1.5,1.75) {Carrier};
	\node[draw=none] at (7.2,1.7) {1};
	\node[draw=none] at (6.6,2.25) {0};
	\node[draw=none] at (7.8,1.15) {0};
\end{tikzpicture}