### 2022-10-13 assets/illustration/ris.tex

\begin{tikzpicture}[font=\LARGE,every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [dart,shape border rotate=90,minimum width=1cm] {};
	\node at (5,3.25) [semicircle,shape border rotate=180,minimum width=1.5cm] {};
	\node at (8,0.35) [dart,shape border rotate=270,minimum width=1cm] {};

	\node[draw=none] at (0,-1) {Transmitter};
	\node[draw=none] at (5,4.4375) {Reflecting\\Element};
	\node[draw=none] at (8,-1) {Receiver};

	\draw[decorate,decoration={waves,segment length=5mm}] (1,-0.25) -- (7.375,-0.25);
	\draw[decorate,decoration={waves,segment length=5mm}] (0.5,0.75) -- (4,3);
	\draw[decorate,decoration={waves,segment length=5mm}] (6,3) -- (8,1.25);

	\node[draw=none] at (4.875,1.125) {Modulated\\Signal};
\end{tikzpicture}