### 2022-10-18 assets/simulation/region_comparison.tex
:bento: Add scattering rates comparison

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.46667,0.67451,0.18824}%
\definecolor{mycolor2}{rgb}{0.49400,0.18400,0.55600}%
\definecolor{mycolor3}{rgb}{0.92900,0.69400,0.12500}%
\definecolor{mycolor4}{rgb}{0.85000,0.32500,0.09800}%
\definecolor{mycolor5}{rgb}{0.00000,0.44700,0.74100}%
%
\begin{tikzpicture}

\begin{axis}[%
width=4.079in,
height=3.432in,
at={(0.684in,0.463in)},
scale only axis,
xmin=0,
xmax=7.08488905081188,
xlabel style={font=\color{white!15!black}},
xlabel={Primary Rate [bits/s/Hz]},
ymin=0,
ymax=2,
ylabel style={font=\color{white!15!black}},
ylabel={Total Backscatter Rate [bits/BSP]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.03,0.03)}, anchor=south west, legend cell align=left, align=left, draw=white!15!black},
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\large}, legend style={font=\large}, reverse legend, every axis plot/.append style={line width=2pt}
]
\addplot [color=mycolor1, mark=triangle, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
6.41670947751651	1.6054968458145\\
0	1.6054968458145\\
0	0\\
7.08488904258997	0\\
7.08488904258997	4.1836174712136e-07\\
6.90245078291004	1.01381402853452\\
6.6407882228124	1.4980015567956\\
6.57239022993099	1.55479524496005\\
6.51932412280019	1.58378871636645\\
6.49722825073873	1.59220191008592\\
6.4775306470264	1.59794430837116\\
6.4598927531162	1.60170353740201\\
6.44402970380102	1.60398330481762\\
6.43668727400095	1.6046886802456\\
6.42970223763566	1.60515550025508\\
6.42700280669212	1.60528273332009\\
6.42435520073694	1.60537879376922\\
6.4217579842616	1.60544542043883\\
6.41920966721981	1.60548425498564\\
6.41670947751651	1.6054968458145\\
};
\addlegendentry{RIScatter}

\addplot[only marks, mark=x, mark options={}, mark size=3.5355pt, draw=mycolor2] table[row sep=crcr]{%
x	y\\
7.08488905081188	0\\
};
\addlegendentry{RIS}

\addplot[only marks, mark=square, mark options={}, mark size=2.5000pt, draw=mycolor3] table[row sep=crcr]{%
x	y\\
6.44440818236646	2\\
};
\addlegendentry{SR}

\addplot[only marks, mark=+, mark options={}, mark size=3.5355pt, draw=mycolor4] table[row sep=crcr]{%
x	y\\
3.85295770698096	1.53088335179488\\
};
\addlegendentry{AmBC}

\addplot[only marks, mark=o, mark options={}, mark size=3.5355pt, draw=mycolor5] table[row sep=crcr]{%
x	y\\
0	1.99999999999997\\
};
\addlegendentry{BBC}

\end{axis}
\end{tikzpicture}%