### 2022-10-28 assets/illustration/time_structure.tex
:bento: Update time structure

\begin{tikzpicture}
	% BC/AmBC/SR
	\foreach \x in {0,...,3}{
			\draw[blue,fill=blue!30!white,postaction={pattern=horizontal lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}
	\foreach \x in {4,...,7}{
			\draw[blue,fill=blue!30!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}
	\foreach \x in {8,...,11}{
			\draw[blue,fill=blue!30!white,postaction={pattern=vertical lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}
	\foreach \x in {12,...,15}{
			\draw[blue,fill=blue!30!white,postaction={pattern=north east lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}

	% RIS
	\foreach \x in {0,...,15}{
			\draw[red,fill=red!30!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,2) rectangle ++(0.5,0.25);
		}

	% RIScatter
	\foreach \x in {0,...,1}{
			\draw[brown,fill=brown!30!white,postaction={pattern=horizontal lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}
	\foreach \x in {2,...,7}{
			\draw[brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}
	\foreach \x in {8,...,11}{
			\draw[brown,fill=brown!30!white,postaction={pattern=vertical lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}
	\foreach \x in {12,...,15}{
			\draw[brown,fill=brown!30!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}

	\foreach \x in {1,...,7}{
			\draw[dotted] (\x,1) -- (\x,3.75);
		}
	\draw[latex-latex] (0,3.5) -- (0.5,3.5) node[above,midway,yshift=1]{PB};
	\draw[latex-latex] (0,4.1) -- (1,4.1) node[above,midway,yshift=1]{BB};
	\draw[latex-latex] (0,4.7) -- (8,4.7) node[above,midway,yshift=1]{Channel Block};

	\node at (-2,3.125) {BackCom/AmBC/SR};
	\node at (-2,2.125) {RIS};
	\node at (-2,1.125) {RIScatter};

	\draw [pattern=horizontal lines](0.15,0) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 1};
	\draw [pattern=north west lines](2.15,0) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 2};
	\draw [pattern=vertical lines](4.15,0) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 3};
	\draw [pattern=north east lines](6.15,0) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 4};
\end{tikzpicture}