### 2022-10-29 assets/simulation/region_comparison.tex

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.46667,0.67451,0.18824}%
\definecolor{mycolor2}{rgb}{0.49400,0.18400,0.55600}%
\definecolor{mycolor3}{rgb}{0.92900,0.69400,0.12500}%
\definecolor{mycolor4}{rgb}{0.85000,0.32500,0.09800}%
\definecolor{mycolor5}{rgb}{0.00000,0.44700,0.74100}%
%
\begin{tikzpicture}

	\begin{axis}[%
			width=4.079in,
			height=1.587in,
			at={(0.684in,0.361in)},
			scale only axis,
			xmin=0,
			xmax=7.08488905081188,
			xlabel style={font=\color{white!15!black}},
			xlabel={Primary Rate [bits/s/Hz]},
			ymin=0,
			ymax=2,
			ylabel style={font=\color{white!15!black}},
			ylabel={Total Backscatter\\Rate [bits/BB]},
			axis background/.style={fill=white},
			xmajorgrids,
			ymajorgrids,
			legend style={at={(0.03,0.03)}, anchor=south west, legend cell align=left, align=left, draw=white!15!black},
			title style={font=\LARGE},
			label style={font=\LARGE},
			ticklabel style={font=\Large},
			legend style={font=\Large},
			reverse legend,
			every axis plot/.append style={line width=2pt},
			align=center
		]
		\addplot [color=mycolor1, line width=2.0pt, mark=triangle, mark options={solid, mycolor1}]
		table[row sep=crcr]{%
				6.41670947751651	1.6054968458145\\
				0	1.6054968458145\\
				0	0\\
				7.08488904258997	0\\
				7.08488904258997	4.1836174712136e-07\\
				6.90245078291004	1.01381402853452\\
				6.6407882228124	1.4980015567956\\
				6.57239022993099	1.55479524496005\\
				6.51932412280019	1.58378871636645\\
				6.49722825073873	1.59220191008592\\
				6.4775306470264	1.59794430837116\\
				6.4598927531162	1.60170353740201\\
				6.44402970380102	1.60398330481762\\
				6.43668727400095	1.6046886802456\\
				6.42970223763566	1.60515550025508\\
				6.42700280669212	1.60528273332009\\
				6.42435520073694	1.60537879376922\\
				6.4217579842616	1.60544542043883\\
				6.41920966721981	1.60548425498564\\
				6.41670947751651	1.6054968458145\\
			};
		\addlegendentry{RIScatter}

		\addplot[only marks, mark=x, mark options={}, mark size=3.5355pt, draw=mycolor2] table[row sep=crcr]{%
				x	y\\
				7.08488905081188	0\\
			};
		\addlegendentry{RIS}

		\addplot[only marks, mark=square, mark options={}, mark size=2.5000pt, draw=mycolor3] table[row sep=crcr]{%
				x	y\\
				6.44440818236646	2\\
			};
		\addlegendentry{SR}

		\addplot[only marks, mark=+, mark options={}, mark size=3.5355pt, draw=mycolor4] table[row sep=crcr]{%
				x	y\\
				3.85295770698096	1.53088335179488\\
			};
		\addlegendentry{AmBC}

		\addplot[only marks, mark=o, mark options={}, mark size=3.5355pt, draw=mycolor5] table[row sep=crcr]{%
				x	y\\
				0	1.99999999999997\\
			};
		\addlegendentry{BBC}

	\end{axis}
\end{tikzpicture}%