### 2023-10-22 poster/custom-defs.tex

% ====================
% BibLaTeX
% ====================

% \usepackage[backend=biber,
% 	bibstyle=authoryear,
% 	citestyle=authoryear,
% 	style=authoryear,
% 	maxcitenames=2,
% 	maxbibnames=20, % limit the length of list of names (authors/editors/etc.)
% 	sorting=ydnt, % sort references by year (descending), name, title
% 	dashed=false, % show authors instead of dash in publications having the same authors
% 	giveninits=true % render authors' given name initials and not the full given names
% ]{biblatex}
% %% Biblatex with Beamer bibliography icons
% \setbeamertemplate{bibliography item}{%
% 	\ifboolexpr{ test {\ifentrytype{book}} or test {\ifentrytype{mvbook}}
% 		or test {\ifentrytype{collection}} or test {\ifentrytype{mvcollection}}
% 		or test {\ifentrytype{reference}} or test {\ifentrytype{mvreference}} }
% 	{\setbeamertemplate{bibliography item}[book]}
% 	{\ifentrytype{online}
% 		{\setbeamertemplate{bibliography item}[online]}
% 		{\setbeamertemplate{bibliography item}[article]}}%
% 	\usebeamertemplate{bibliography item}}
% \defbibenvironment{bibliography}
% {\list{}
% 	{\settowidth{\labelwidth}{\usebeamertemplate{bibliography item}}%
% 		\setlength{\leftmargin}{\labelwidth}%
% 		\setlength{\labelsep}{\biblabelsep}%
% 		\addtolength{\leftmargin}{\labelsep}%
% 		\setlength{\itemsep}{\bibitemsep}%
% 		\setlength{\parsep}{\bibparsep}}}
% {\endlist}
% {\item}
% %% Redefine \refname
% \renewcommand{\bibname}{References}
% %% Redefine \parencite to use square brackets instead of braces
% \DeclareCiteCommand{\parencite}
% {\usebibmacro{prenote}}
% {\usebibmacro{citeindex}%
% 	\printtext[bibhyperref]{[\usebibmacro{cite}]}}
% {\multicitedelim}
% {\usebibmacro{postnote}}
% %% Highlight author names using Beamer data annotation
% %% Usage: add a new line `author+an = {<author-order>=highlight}` to an entry
% %% For example: author+an = {3=highlight} => highlight the 3rd author name
% \AtBeginBibliography{
% 	\renewcommand*{\mkbibnamegiven}[1]{%
% 		\ifitemannotation{highlight}
% 		{\textbf{#1}}
% 		{#1}%
% 	}

% 	\renewcommand*{\mkbibnamefamily}[1]{%
% 		\ifitemannotation{highlight}
% 		{\textbf{#1}}
% 		{#1}%
% 	}
% }

% ====================
% Boxes with tcolorbox
% ====================
\usepackage[most]{tcolorbox}

%%% Beamer colors in boxes

\newcommand{\beamercolorsinboxes}[1]{
	\setbeamercolor{itemize item}{fg=#1!75!black}
	\setbeamercolor{itemize/enumerate body}{fg=#1!65!white}
	\setbeamercolor{itemize/enumerate subbody}{fg=#1!65!white}
	\setbeamercolor{item projected}{fg=white, bg=#1!75!black}
}

%%% Highlight Oval Box
\newtcbox{\xmybox}[1][red]{on line,
	arc=7pt,colback=#1!10!white,colframe=#1!50!black,
	before upper={\rule[-3pt]{0pt}{10pt}},boxrule=1pt,
	boxsep=0pt,left=6pt,right=6pt,top=2pt,bottom=2pt}
%%% Box for stating problems
%%%%%%%%
%Usage: (similar for infobox)
%	\begin{defbox}{title}
%		contents
%	\end{defbox}
%%%%%%%%
\newtcolorbox{defbox}[1]{%
	enhanced,
	attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	colback=cyan!5!white,
	colframe=cyan!75!black,
	coltitle=cyan!80!black,
	%	left=0mm,right=0mm,top=2mm,bottom=0mm,
	title={#1},
	fonttitle=\bfseries\large, fontupper=\color{cyan!65!white},
	boxed title style={colback=cyan!5!white,colframe=cyan!75!black},
	before upper={
			\beamercolorsinboxes{cyan}
		}
}%
%%% Box for announcement
\newtcolorbox{infobox}[1]{%
	enhanced,
	attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	colback=yellow,
	colframe=red!75!black,
	coltitle=red!75!black,
	%	left=0mm,right=0mm,top=2mm,bottom=0mm,
	title={#1},
	fonttitle=\bfseries\large, fontupper=\color{red!65!white},
	boxed title style={colback=yellow,colframe=red!75!black},
	before upper={
			\beamercolorsinboxes{red}
		}
}%
%%% Box for example
\newtcolorbox{exabox}[1]{%
	enhanced,
	attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	colframe=brown!75!black,colback=brown!5!white,coltitle=brown!50!brown!75!black,
	%	left=0mm,right=0mm,top=2mm,bottom=0mm,
	title={#1},
	fonttitle=\bfseries\large, fontupper=\color{brown!65!white},
	boxed title style={colback=brown!5!white,coltitle=brown!50!brown!75!black},
	before upper={
			\beamercolorsinboxes{brown}
		}
}%
%%% Theorem Box
%%%%%%%%
%Usage: (similar for conjecture, lemma, etc.)
%	\begin{thm}{title}{nameref}
%		contents
%	\end{thm}
% Use \ref{thm:nameref} to refer to the theorem
%%%%%%%%
%%%% Use \newtcbtheorem[number within=section]{thm} to number within each section
\newtcbtheorem[]{thm}%
% {Theorem}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
{Block}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	enhanced jigsaw,
	top=2mm,bottom=6mm,left=0mm,right=0mm,
	% fonttitle=\bfseries\large,fontupper=\itshape\color{blue!65!white},
	fonttitle=\bfseries\large,fontupper=\color{blue!65!white},
	colframe=blue!75!black,colback=blue!5!white,coltitle=blue!50!blue!75!black,
	boxed title style={colback=blue!5!white,coltitle=blue!50!blue!75!black},
	before upper={
			\beamercolorsinboxes{blue}
		}
}{thm}%
%%% Proposition Box
\newtcbtheorem[use counter from=thm]{prop}%
{Proposition}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	enhanced jigsaw,
	%	top=2mm,bottom=0mm,left=0mm,right=0mm,
	fonttitle=\bfseries\large,fontupper=\itshape,
	colframe=gray!75!black,colback=gray!5!white,coltitle=gray!50!gray!75!black,
	boxed title style={colback=gray!5!white,coltitle=gray!50!gray!75!black},
	before upper={
			\beamercolorsinboxes{gray}
		}
}{prop}%
%%% Conjecture Box
\newtcbtheorem[use counter from=thm]{conj}%
{Conjecture}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	enhanced jigsaw,
	%	top=2mm,bottom=0mm,left=0mm,right=0mm,
	fonttitle=\bfseries\large,fontupper=\slshape,
	colframe=orange!75!black,colback=orange!5!white,coltitle=orange!50!orange!75!black,
	boxed title style={colback=orange!5!white,coltitle=orange!50!orange!75!black},
	before upper={
			\beamercolorsinboxes{orange}
		}
}{conj}%
%%% Lemma Box
\newtcbtheorem[use counter from=thm]{lem}%
{Lemma}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	enhanced jigsaw,
	%	top=2mm,bottom=0mm,left=0mm,right=0mm,
	fonttitle=\bfseries\large,fontupper=\itshape,
	colframe=green!75!black,colback=green!5!white,coltitle=green!50!green!75!black,
	boxed title style={colback=green!5!white,coltitle=green!50!green!75!black},
	before upper={
			\beamercolorsinboxes{green}
		}
}{lem}%
%%% Claim Box
\newtcbtheorem[use counter from=thm]{clm}%
{Claim}{attach boxed title to top 	left={xshift=5mm,yshift=-5mm,yshifttext=-5mm},
	enhanced jigsaw,
	%	top=2mm,bottom=0mm,left=0mm,right=0mm,
	fonttitle=\bfseries\large,fontupper=\itshape,
	colframe=pink!75!black,colback=pink!5!white,coltitle=pink!50!pink!75!black,
	boxed title style={colback=pink!5!white,coltitle=pink!50!pink!75!black},
	before upper={
			\beamercolorsinboxes{pink}
		}
}{clm}%